---
title: "PA1c"
author: "Yan Feng"
date: "June 27, 2017"
output: md_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
### loading
```{r loading}
activity <- read.csv("activity.csv", na.strings = NA, stringsAsFactors = F, colClasses = c("numeric", "character", "numeric"))
```
### preprocessing
```{r preprocessing}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```
## What is mean total number of steps taken per day?
### total number per day
```{r total steps}
stepsperday <- sapply(split(activity$steps, activity$date), sum, na.rm = T)
```
### histogram and barplot
```{r histogram and barplot}
hist(stepsperday, main = "Total number of steps per day", 
     xlab = "total steps per day")
barplot(stepsperday, main = "Total number of steps per day", 
        ylab = "total steps per day")
```

### report mean and median
```{r mean and median}
sprintf("The mean of total number of steps per day is %.f.", mean(stepsperday))
sprintf("The median of total number of steps per day is %.f.",
        median(stepsperday))
```

## What is the average daily activity pattern?
### time series plot
```{r time seris plot}
acrossallday <- do.call(cbind, split(activity$steps, activity$date))
acrossallday_interval <- apply(acrossallday, 1, mean, na.rm = TRUE)
interval <- unique(as.character(activity$interval))
names(acrossallday_interval) <- interval
plot(interval, acrossallday_interval, type = 'l',
     ylab = "average steps per interval",
     xlab = "interval")
```

### Which 5-minute interval consists the max steps?
```{r which interval max step}
max_interval <- names(acrossallday_interval[which.max(acrossallday_interval)])
sprintf(paste(max_interval, "interval consists the max steps."))
```

## Imputing missing values
### total number of missing values
```{r total number of NA}
sum_na <- sum(is.na(activity$steps))
sprintf("The total number of missing values is %.f", sum_na)
```

### Missing values filled in
```{r fill in missing values}
# Using mean for that interval to fill in
patch <- function(a) {
        mask <- is.na(a)
        a[mask] <- acrossallday_interval[mask]
        a
}
activity_1 <- transform(activity, interval = as.character(activity$interval))
stepsbydaylist <- split(activity_1$steps, activity_1$date)
patchsteps <- lapply(stepsbydaylist, patch)
activity_1$steps <- unlist(patchsteps)
```

### total number per day after patching
```{r total steps after patching}
patchstepsperday <- sapply(split(activity_1$steps, activity$date), sum)
```

### histogram after patching
```{r histogram}
hist(patchstepsperday, main = "Total number of steps per day", 
     xlab = "total steps per day")
```

### report mean and median
```{r mean and median per day}
sprintf("The mean of total number of steps per day is %.f.", 
        mean(patchstepsperday))
sprintf("The median of total number of steps per day is %.f.",
        median(patchstepsperday))
```

## Are there differences in activity patterns between weekdays and weekends?
### create new variable weekday
```{r create new variable weekday}
weekends <- c("Saturday", "Sunday")
library(dplyr)
activity <- mutate(activity, weekday = ifelse(weekdays(activity$date) %in% weekends, "weekend", "weekday"))
activity <- mutate(activity, weekday = as.factor(weekday))
activity <- mutate(activity, interval = as.character(interval))
```

### make plot of weekdays and weekends
```{r make plots of weekdays and weekends}
activity_weekday <- subset(activity, weekday == "weekday")
activity_weekend <- subset(activity, weekday == "weekend")
interval_num <- as.numeric(interval)
interval_wd <- data.frame(split(activity_weekday$steps, activity_weekday$date))
interval_we <- data.frame(split(activity_weekend$steps, activity_weekend$date))
interval_wd_day <- rowMeans(interval_wd, na.rm = T)
interval_we_day <- rowMeans(interval_we, na.rm = T)
interval_wd_day_df <- data.frame(interval_wd_day, weekday = "weekday", interval = interval_num)
interval_we_day_df <- data.frame(interval_we_day, weekday = "weekend", interval = interval_num)
names(interval_wd_day_df)[1] <- "steps"
names(interval_we_day_df)[1] <- "steps"
interval_wd_we_df <- rbind(interval_wd_day_df, interval_we_day_df)
library(lattice)
xyplot(steps ~ interval | weekday, data = interval_wd_we_df, layout = c(1, 2), type = 'l')
```









